### 2016-02-22 ch02-qlogms/excitation-design.tex
Many small changes to CH02, CH05

\section{Introduction}
\label{sec:excitation:intro}
Input design deals with the task of choosing an excitation signal in a way so that the information obtained from the experiment is maximized. 
This is a problem which has a long history in statistics. 
In system identification, the focus has mostly been on input design for dynamic systems.
A common approach is to design the input spectrum and thereby maximizing the information matrix of the estimates, see e.g. \cite{Fedorov1972,Goodwin1977}.

A complicating issue in identification of dynamic system is the fact that the information matrix often depends on the true parameter values. 
Consequently, an optimal input design, based on the information matrix, will depend on quantities that are not know at the time of the design. 
There are two main solutions to this discussed in the engineering literature:
\begin{enumerate}
\item Iterative input design where the design updated as more information about the system is available, see e.g. \cite{Hjalmarsson2005,Gevers2005}.
\item Robust input design one tries to make the input design robust to uncertainty in the a priori knowledge of the true parameter values, see e.g. \cite{Rojas2007} and references therein. 
\end{enumerate}

In this chapter we focus on robust excitation and we address the situation where there is very limited a priori information about the system to be identified. 
We will consider input design when it is known that the system has resonances, possibly spanning a large frequency band. 
It is however not necessary to know how many resonances are in the system. 
The advantage of the proposed input design is that regardless of where these resonances are located, the resulting transfer function variance will be the same, for a given damping in the system.

Experiment design is the first step in the system identification loop, as illustrated in \figref{fig:intro:identification-cycle}.
Consequently, the design that is chosen influences every further step in the identification process and limit the model quality that can be attained.

\subsection{Approach in this chapter}
\label{sec:excitation:intro:approach}
In this chapter, the goal is to construct excitation signals that bridge the properties of generic and robust excitation signals to alleviate the chicken-and-the-egg problem that optimal strategies typically incur.
In particular, a generic excitation signal will be used but its properties are tuned in such a way that it attains some properties of robust excitation signals without the requirement of very specific prior information.

\input{ch02-qlogms/logMS_IMTC.tex}

\section{Introduction}
  It often occurs that we wish to measure the \gls{FRF} of a \gls{LTI} system in a frequency band that covers several decades~\citep{Bragos2001,Sanchez2011,Niedostatkiewicz2009,VanGheem2004}.
  Of course, we also want to attain a specified uncertainty level over the full frequency span.

@@

\section{Input design}\label{sec:excitation:inputDesign}
For the theoretical analysis, we consider SISO output error systems, i.e.
\begin{align}
y(t) &=G(s,\theta)u(t)  + e(t),
\end{align}
where $u(t)$ is a known input sequence, $y(t)$ the output sequence and $e(t)$ is zero mean Gaussian white noise.
$G$ is a rational, time invariant transfer function in continuous time. 
For notational simplicity, $s$ is used both as the complex Laplace variable and the differentiation operator; it should be clear from the context which is used. 
The transfer function is parameterized by the vector $\theta$.
These parameters are unknown and therefore need to be estimated.

@@

  \subsection{The system under consideration}
  We consider the transfer function models of generic \gls{LTI} systems of arbitrary but finite orders $n_b / n_a$:
  \begin{align}
    G\left( \Omega, \theta \right) = \frac{B\left( \Omega,\theta \right)}
                                          {A\left( \Omega,\theta \right)}
                                   = \frac{\sum\limits_{i=0}^{n_b-1} b_i \Omega^i}
                                          {\sum\limits_{i=0}^{n_a-1} a_i \Omega^i}
    \text{.} \label{eq:model}
  \end{align}

@@

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion:
  \TODO{move into main text}
Remember that the rational transfer function of any strictly proper \gls{LTI} system $G$ can be decomposed in its partial fraction form:
\begin{align}
  G(s) &= 
    \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\sum\limits_{i=0}^{\numel{A}} a_i s^i} 
        =    
   \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\prod\limits_{i=1}^{I_1} \left( s - \pole[i] \right)^{M_{1,i}} 
               \prod\limits_{i=1}^{I_2} \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{M_{2,i}}}
        \\
        &  = 
        \sum_{i=1}^{I_1} 
          \sum_{\multiplicity=1}^{M_{1,i}} 
            \frac{\dc[i]}
                      { \left( s - \pole[i] \right)^{\multiplicity}}
            + 
        \sum_{i=1}^{I_2} 
          \sum_{\multiplicity=1}^{M_{2,i}}
           \frac{\dc[i,\multiplicity] \wn[i]^{2\multiplicity} \left( s + \wz[i] \right)}
                     {\wz[i] \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{\multiplicity}}
        \label{eq:excitation:PFE}
\end{align}
where $I_1$ denotes the number of real-valued poles, $I_2$ denotes the number of complex conjugated pole pairs.
$M_{1,i}$ and $M_{2,i}$ denotes the multiplicity of the corresponding poles, such that $\numel{A} = \sum_i^{I_1} M_{1,i} + 2 \sum_{i}^{I_2} M_{2,i}$.
Each term in \eqref{eq:excitation:PFE} represents a \term{sub-system} of the system $G(s)$.  
  Within these sub-systems, we can distinguish first-order subsystems and second-order subsystems, the latter of which describe resonant phenomena.
  Since such resonances are often a limiting factor, we first focus on these resonant sub-systems.

@@

\subsection{Multisine Excitations}
  To excite the system under test we consider a generalization of multisine excitations \citep{Pintelon2012}:
  \begin{definition}[Generalized multisine] \label{def:generalized-MS}
  A generalized multisine $u(t)$ is a signal consisting of $F$ sine waves with a different frequencies $f_k$, amplitudes $A_k$ and phase shifts $\phi_k$:
  \begin{equation}
    u \left( t\right) = \frac{1}{\sqrt{F}}
   \sum_{k=1}^{F} 
     A_k 
     \sin 
       \left(2\pi f_k t + \phi_k \right)
  \text{.}
  \label{eq:MultisineCT}
  \end{equation}
  For simplicity, one can stipulate that the frequency grid is sorted, i.e. $f_1 < f_2 < \ldots < f_F$.
  \end{definition}

  We limit ourselves, without loss of generality, to random phase multisines.
  Hence, $\phi_k$ is the outcome of a uniform random process over $\left[0,2\pi\right[$.
  This leaves two sets of free parameters to create a signal that suits our needs:
    the amplitude of the spectral lines $A_k$ and
    the frequency grid $\left\{f_1, \ldots, f_F \right\}$.
  Note that the $k^{\text{th}}$ term in \eqref{eq:MultisineCT} is has a period $T_k \isdef f_k^{-1}$.

@@

  Such an equidistant grid is a very common choice.
  Note that it has more frequency lines per decade (or octave) at higher frequencies.
  Therefore, the excitation power per octave is larger at higher frequencies
  when a constant amplitude spectrum ($A_k = A_{\mathrm{in}}$) is used.
  For a constant amplitude spectrum, this means the dynamics that are present at lower frequencies are plagued by a higher uncertainty than the ones located at higher frequencies.

@@

  For that reason, a logarithmic multisine is a good candidate for our purposes.
  However, it is more involved both to generate and to analyze such a signal, as the excited frequencies don't lie on a commensurate frequency grid.
  Especially for periodic measurements, this is a major downside since the period of such signal does not always exist (i.e. this leads to infinite measurement times).
  It is therefore not practical to use a perfectly logarithmically spaced frequency grid as the \gls{DFT} or \gls{CZT}~\citep{Rabiner2004} explicitly rely on a commensurate frequency grid.

>>>>

\section{Introduction}
\label{sec:excitation:intro}
Input design deals with the task of choosing an excitation signal in a way so that the information obtained from the experiment is maximized. 
This is a problem which has a long history in statistics. 
In system identification, the focus has mostly been on input design for dynamic systems.
A common approach is to design the input spectrum and thereby maximizing the information matrix of the estimates, see e.g. \citep{Fedorov1972,Goodwin1977}.

A complicating issue in identification of dynamic system is the fact that the information matrix often depends on the true parameter values. 
Consequently, an optimal input design, based on the information matrix, will depend on quantities that are not know at the time of the design.
There are two main solutions to this in the engineering literature:
\begin{enumerate}
\item Iterative input design where the design updated as more information about the system is available, see e.g. \citep{Hjalmarsson2005,Gevers2005}.
The disadvantage of this approach is that multiple experiments are required to converge to the optimal design.
Moreover, since the signal is optimized for a specific system model, it is likely that changes in the dynamics might be hard to observe using such a signal.

\item Robust input design one tries to make the input design robust to uncertainty in the prior knowledge of the parameter values, see e.g. \citep{Rojas2007} and references therein. 
While the signal is more robust towards uncertainty in the prior knowledge, it requires more forethought and experimentation than (nominal) optimal designs.
E.g. the approach followed by \citet{Rojas2007} relies on both nominal parameter values and the distribution function of these parameters to design a robust (optimal) excitation signal.
\end{enumerate}

\subsection{Approach in this chapter}
\label{sec:excitation:intro:approach}
In this chapter, we focus on robust excitation and we address the situation where there is very limited a priori information about the system to be identified. 
We will consider input design when it is known that the system has resonances, possibly spanning a frequency band that may cover several decades. 
It is however not necessary to know how many resonances are in the system and at what precise frequencies those may occur.
Such problem settings occur in many diverse fields of experimentation, such as the measurement of biochemical impedances\citep{Bragos2001,Sanchez2011}, electrochemical impedance spectroscopy~\citep{Niedostatkiewicz2009,VanGheem2004}, electronics\TODO{citation}, mechatronics\TODO{citation} and vibration testing\TODO{citation (ISMA?)}.

The advantage of the proposed input design is that regardless of where these resonances are located, the resulting transfer function variance will be the same, for a given damping in the system.

In this chapter, the goal is to construct excitation signals that bridge the properties of generic and robust excitation signals to alleviate the chicken-and-the-egg problem that optimal strategies typically incur.
In particular, a generic excitation signal will be used, but its properties are tuned in such a way that it attains some properties of robust excitation signals without the requirement of very specific prior information.

\section{Introduction}
  It often occurs that we wish to measure the \gls{FRF} of a \gls{LTI} system in a frequency band that covers several decades~.
  Of course, we also want to attain a specified uncertainty level over the full frequency span.

@@

\section{Input design}
\label{sec:excitation:inputDesign}
For the theoretical analysis, we consider \gls{SISO} \gls{OE} systems, i.e.
\begin{align}
y(t) &=G(s,\theta)u(t)  + e(t),
\end{align}
where $u(t)$ is a known input sequence, $y(t)$ the output sequence and $e(t)$ is zero mean Gaussian white noise.
$G$ is a rational, time invariant transfer function in continuous time. 
For notational simplicity, $s$ is used both as the complex Laplace variable and the differentiation operator; it should be clear from the context which is used. 
The transfer function is parameterized by the vector $\theta$.
These parameters are unknown and therefore need to be estimated.

@@

  \subsection{The system under consideration}
  We consider the transfer function models of generic \gls{LTI} systems of arbitrary but finite orders $\order{B} / \order{A}$:
  \begin{align}
    G\left( \Omega, \theta \right) 
    = \frac{B\left( \Omega,\theta \right)}
                  {A\left( \Omega,\theta \right)}
    = \frac{\sum\limits_{i=0}^{\order{B}} b_i \Omega^i}
                  {\sum\limits_{i=0}^{\order{A}} a_i \Omega^i}
    \text{.} \label{eq:model}
  \end{align}

@@

  Such a system can be decomposed into a series of first and second order systems by the partial fraction expansion:
Remember that the rational transfer function of any strictly proper \gls{LTI} system $G$ can be decomposed in its partial fraction form:
\begin{align}
  G(s) &= 
    \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\sum\limits_{i=0}^{\numel{A}} a_i s^i} 
        =    
   \frac{\sum\limits_{i=0}^{\numel{B}} b_i s^i}
              {\prod\limits_{i=1}^{I_1} \left( s - \pole[i] \right)^{M_{1,i}} 
               \prod\limits_{i=1}^{I_2} \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{M_{2,i}}}
        \\
        &  = 
        \sum_{i=1}^{I_1} 
          \sum_{\multiplicity=1}^{M_{1,i}} 
            \frac{\dc[i]}
                      { \left( s - \pole[i] \right)^{\multiplicity}}
            + 
        \sum_{i=1}^{I_2} 
          \sum_{\multiplicity=1}^{M_{2,i}}
           \frac{\dc[i,\multiplicity] \wn[i]^{2\multiplicity} \left( s + \wz[i] \right)}
                     {\wz[i] \left(s^2 + 2 \damping[i] \wn[i] s + \wn[i]^2 \right)^{\multiplicity}}
        \label{eq:excitation:PFE}
\end{align}
where $I_1$ denotes the number of real-valued poles, $I_2$ denotes the number of complex conjugated pole pairs.
$M_{1,i}$ and $M_{2,i}$ denotes the multiplicity of the corresponding poles, such that $\numel{A} = \sum_i^{I_1} M_{1,i} + 2 \sum_{i}^{I_2} M_{2,i}$.
Each term in \eqref{eq:excitation:PFE} represents a \term{sub-system} of the system $G(s)$.  
  Within these sub-systems, we can distinguish first-order subsystems and second-order subsystems, the latter of which describe resonant phenomena.
  Since such resonances are often a limiting factor, we first focus on these resonant sub-systems.

@@

\subsection{Multisine Excitations}
  To excite the system under test we consider a generalization of multisine excitations \citep{Pintelon2012}:
  \begin{definition}[Generalized multisine] \label{def:generalized-MS}
  A generalized multisine $u(t)$ is a signal consisting of $F$ sine waves with a different frequencies $f_k$, amplitudes $A_k$ and phase shifts $\phi_k$:
  \begin{equation}
    u \left( t\right) = \frac{1}{\sqrt{F}}
   \sum_{k=1}^{F} 
     A_k 
     \sin 
       \left(2\pi f_k t + \phi_k \right)
  \text{.}
  \label{eq:MultisineCT}
  \end{equation}
  Without loss of generality, one can stipulate that the frequency grid is sorted, i.e. $f_1 < f_2 < \ldots < f_F$.
  \end{definition}

  For \gls{LTI} systems, it is well-known \TODO{refer to eqn. from Larsson} that the information matrix only depends upon the amplitude spectrum of the input and not upon its phase spectrum.
  As such, in this chapter assume random-phase multisines since this is a reasonable choice for systems that are dominantly linear~\TODO{referentie}.
  Hence, $\phi_k$ is the outcome of a uniform random process over $\left[0,2\pi\right[$.
  This leaves two sets of free parameters to create a signal that suits our needs:
    the amplitude of the spectral lines $A_k$ and
    the frequency grid $\left\{f_1, \ldots, f_F \right\}$.
  Note that the $k^{\text{th}}$ term in \eqref{eq:MultisineCT} is periodic with period $T_k \isdef f_k^{-1}$.

@@

  Such an equidistant grid is a very common choice.
  Note that it has more frequency lines per decade (or octave) at higher frequencies.
  Therefore, the excitation power per octave is larger at higher frequencies
  when a constant amplitude spectrum ($A_k = A_{\mathrm{in}}$) is used.
  This is a rather common choice in multisine excitations when no other constraints are taken into account.
  For a constant amplitude spectrum, this means that a model of the low-frequency dynamics is typically plagued by a higher uncertainty than for models of high-frequency dynamics.
  This is caused by the fact that the lower frequency bands receive less signal power than the high frequency bands.

@@

  For that reason, a logarithmic multisine is a good candidate for our purposes.
  However, it is more involved both to generate and to analyze such a signal, as the excited frequencies do not lie on a commensurate frequency grid.
  Especially for periodic measurements, this is a major downside since the period of such signal does not always exist (i.e. this leads to infinite measurement times).
  It is therefore not practical to use a perfectly logarithmically spaced frequency grid as the \gls{DFT} or \gls{CZT}~\citep{Rabiner2004} explicitly rely on a commensurate frequency grid.
